---
title: 'STP 598: Homework 3'
author: "Antonio Campbell, Sinta Sulistyo, Atta Ullah, Penny Wu"
date: "3/9/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
#=============================================
## Clear .Rdata
rm(list=ls())

## Packages
pkg.list <-c("ggplot2", "reshape2", "knitr","matrixcalc",
             "rootSolve", "Matrix", "kknn", "Metrics", "pracma"
             ,"glmnet")

#install.packages(pkg.list)
lapply(pkg.list, require, character.only = TRUE)

knitr::opts_chunk$set(echo = FALSE, fig.align='center')
#=============================================
```

#1. Problem 1. Basic Optimization, MLE for IID Poisson Data
Suppose $y_i$ is a count then a very common model is to assume the Poisson disttribuion: $$P(Y=y \;|\; \lambda) = \frac{e^{-\lambda} \, \lambda^y}{y!}, \; y = 0,1,2,\ldots$$
Given $Y_i \sim Poisson(\lambda)$ iid, (that is, \(Y_i = y_i\)), what is the MLE of $\lambda$?

If the random variables are iid, then the joint distribution is the product of the marginal distributions. Let $\mathbf{y}$ be the sample $y_1, \dots, y_n$ 
$$L(\lambda \vert \mathbf{y}) = \prod_{i = 1}^{n}\frac{e^{-\lambda} \, \lambda^{y_i}}{y_i!}.$$ We take the log of the likelihood so that we can maximize easily. 
$$\log(L(\lambda \vert \mathbf{y})) = \sum_{i = 1}^{n}\log\bigg(\frac{e^{-\lambda} \, \lambda^{y_i}}{y_i!}\bigg) = -n\lambda -\sum{\log(y_i!)}+ \sum(y_i)\log(\lambda).$$
Differentiate with respect to lambda and set to zero to maximize: $$\frac{\delta \log(L(\lambda \vert \mathbf{y}))}{\delta d\lambda}=-n + \frac{\sum y_i}{\lambda} = 0 \implies \hat\lambda =\frac{\sum y_i}{n}$$. 

The MLE for the parameter of the $poisson$ distribution is the sample mean. 


#Problem 2. Constrained Optimization, Minimum Variance Portfolio

Suppose we are considering investing in $p$ stocks where the uncertain return on the $i$th
stock is denoted by $R_i$, $i=1,2,\dots,p$. Let $R=(R_1,R_2,\dots,R_p)^{'}.$ A portfolio is a given by $w=(w_1,w_2,\dots,w_p)^{'}$ where $w_i$ is the fraction of wealth invested in asset $i$. The $\{w_i\}$
must satisfy $\sum w_i=1$. The return on the portfolio is then $$P = w'R = \sum w_i R_i.$$ We want to find the global \textbf{minimum variance portfolio}: $$\underset{w}{\min} \, Var(P), \;\; \text{subject to} \sum w_i = 1.$$ If we let $\iota=(1,1,\dots,1)^{'}$, the vector of ones, and $Var(R)=\Sigma$ then our problem is $$\underset{w}{\min }\,  w^{'}\Sigma w  \;\; \text{subject to } w^{'}\iota = 1.$$ Find the global minimum variance portfolio in terms of $\Sigma$
and $\iota$.

If we know that $\sum {w_i} = 1$ we can easily say $\sum {w_i}-1 = 0$ or $w^{'}\iota-1 = 0$. We use the Karush-Kuhn-Tucker conditions to minimize. By this we take the derivative with respect to $w$, $$\frac{\delta  L(\lambda,w)}{\delta w} = \frac{\delta  (w^{'}\Sigma{w})}{\delta w} + \frac{\delta  \lambda(w^{'}\iota-1)}{\delta w} = 2\Sigma w + \lambda\iota$$ $$ \implies {\hat w} = -\lambda\Sigma^{-1}\iota/2.$$ To express this solution in terms of $\Sigma, \iota$ we solve for $\lambda$. Premultiply by $\iota^{'}$ and then solve. Since we have that $\iota^{'}{w} = 1$, $$ \iota^{'}{w} = -\lambda\iota^{'}\Sigma^{-1}\iota/2  \implies \lambda = -2/(\iota^{'}\Sigma^{-1}{\iota}).$$ We plug this into our solution of $\hat w$ to have:

$$\hat{w} = \frac{-\lambda\Sigma^{-1}\iota/2}{-2/\iota^{'}\Sigma^{-1}{\iota}} =\frac{\lambda\Sigma^{-1}\iota}{\iota^{'}\Sigma^{-1}{\iota}}$$

$\hat w$ is the $w$ such that $Var(w^{'}R) = Var(P)$ is minimized.


# Problem 3. Polynomial Regression

A basic idea in nonlinear regression is to use polynomial terms.

With one $x$ variable, this means we consider the models: $$Y_i=\beta_0+\beta_1 x_i+\beta_2 x^2_i+\ldots+\beta_p x^p_i+\epsilon_i.$$

Using the simple used cars data (with $n=1,000$) with $Y$= price and $x$=mileage, find the best choice of $p$. 
* Use BIC to pick $p$
* use an out-of-sample criteria to pick $p$

Fit your chosen polynomial mode using all the data and plot the fit on top of the data. Do you like it? Also plot the fits for a $p$ that is “way to big”. What is wrong with it?

```{r echo=false,cache=true}

cars = read.csv("http://www.rob-mcculloch.org/data/susedcars.csv", header = TRUE)
head(cars)
```

First, we need to create a dataframe for $x$ varaible:
```{r echo=false,cache=true}
attach(cars)
#price = price/1000
mileage = mileage/1000
ddf<- data.frame(price=price,xi = mileage, xi2 = mileage^2,xi3 = mileage^3,
               xi4 = mileage^4, xi5=mileage^5,xi6=mileage^6, xi7=mileage^7,
               xi8=mileage^8,xi9=mileage^9,xi10=mileage^10)
head(ddf,3)
detach(cars)
attach(ddf)
```   

Now we will apply forward selection with BIC to choose the best $p$.
## Forward selection with BIC


```{r warning=false,cache=true}
source("do-stepcv.R")
n <- nrow(ddf)
##do forwards stepAIC, stopping when adding any var makes BIC bigger.
nullmod = lm(price~1,ddf) #null model, no variables
lmtran =lm(price~.,ddf) #full model
#bictran = extractAIC(lmtran,k=log(n))
#print(summary(lmtran))
#bictran


fwd = stepAIC(nullmod,scope=formula(lmtran),direction="forward",k=log(n),trace=0)
print(summary(fwd))
bicfwd = extractAIC(fwd,k=log(n))
cat("bic for forward model is: ",bicfwd,"\n")

```



```{r }
source("do-stepcv.R")
yind=1
xind=2:ncol(ddf)
set.seed(199)
folds = getfolds(5,nrow(ddf))
nstep = 7

stcv = stepcv(ddf,yind,xind,formula(lmtran),folds,nstep)

rmse=sqrt(apply(stcv,1,sum)/n)
plot(0:nstep,rmse,xlab="num var",ylab="RMSE",
      cex.lab=1.5,cex.axis=1.5,type="b",col="red",pch=16)


```






